---
title: "Overview of gameday"
author: "Wenyi Xia"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

## Overview of gameday

Living in a country with so many hockey lovers, you may also have the urge to check:

* Whether you NHL team has a gameday on a particular day?
* What's the game results of the NHL teams on a particular day?

The gameday package could make these checks easy and fast.

The goal of this document is to introduce you to the basic funcitons that gameday provides.

## Install and Load the package

Install `gameday` from GitHub:

```{r eval = FALSE}
devtools::install_github("wenyixia/gameday")
```

Load the package.

```{r}
library(gameday)
```

## Play with the function `gday()`

Let's play with the `gday()` function.

After you install and load the `gameday` package, in your `R Console`, you could type `?gday` to read the documentation of this function. There are two arguments in the `gday()` function. They are `team` and `date`.

The default of `gday()` is `team = "canucks"` and `date = Sys.Date()`.

```{r}
gday()
```

Now you could go to the [NHL webpage](http://www.nhl.com/) to check the information, which must be consistent with the result returned by `gday()`.

Let's then play with the `team` argument.

```{r}
gday("Predators")
gday("Nashville")
```

Team names and home city names return this same result.

What if the city does not have a NHL team or you make a typo in the name? It will return a error specifying the problem.

```{r}
#gday("Seattle")
```

How do I know that the `team` input is valid or not? Thanks to the data `arena_team` in this package. `arena_team` gives NHL Team Names with the corresponding home city and arena. Type `?arena_team` in `R Console` and you will see the detailed documentation.

Cases are ignored in the fucntion.

```{r}
gday("VANCOUVER")
gday("vancouver")
```

The default of argument `date` is today. Let's then play with the `date` argument.

```{r}
gday("Flames", "2014-11-16")
```

What if the format of the `date` input is wrong? It will tell you the problem.

```{r}
#gday("Flames", "201-411-116")
```

## Play with the function `scores()`

As you type `?scores` in `R console`, you will know that `scores()` returns a data.frame about NHL game information on a particular day.

The default value is today.

```{r}
scores()
```

If you want to know the game infomation on a particular day, then type in the date.

```{r}
scores("2014-11-19")
```

If your input date is an invalid date, it will give you an error.

```{r}
#scores("201-411-19")
```

Lastly, cut off your internet connection and call `gday()` or `scores()`. Both function will throw you an error telling you that there is no internet connnection.

Note that in the above examples, I used `#` for functions that would return an error because otherwise when I `Build > Check`, there would be errors for sourcing `overview.R`.

## How this package was made

The [Class Activity 1](http://stat545-ubc.github.io/packages02_activity.html) and [Class Activity 2](http://stat545-ubc.github.io/packages03_activity_part2.html#vignette) lead us to a package that is half-made. Based on it, I recorded my workflow below.

#### Helper function `Internet_connection()`
Following the instructions of the [Homework](http://stat545-ubc.github.io/hw10_package.html), I first tried to add the function `Internet_connection()`. I copied the function that is already provided in the homework.

```
internet_connection <- function() {
  tryCatch({RCurl::getURL("www.google.com"); TRUE},
    error = function(err) FALSE)
}
```
Thanks for our TAs. Otherwise I would have to spend much time to figure out this function by myself.

Initially, I copied this function directly to the R script `gday.R`. When I `build & Reload`, the package did not know which function to export. Then I googled the problem, and found out that the helper functions could be created in a different R script. Even though the functions are in different R script, the exported function will know where to call the helper function. R package turns out to be much more smart than I thought.

Therefore, I added the R script of `Internet_connection()` to the `/R` folder. As I did not add any documentation such as `@export`, the package will not export this function. Then I added the `if()` and `stop()` statements that we learned from [Write your own function](http://stat545-ubc.github.io/block011_write-your-own-function-01.html) as follows to the `gday.r`

```
	if(internet_connection() == FALSE) {
		stop('Sorry, unable to connect to the Internet.
				    Please check your internet connection.')
	}
```
Lastly, I `build & Reload`, cut off my internet connection to see if this function works, and it worked.

#### Helper function `check_date()`
I followed the template of `Internet_connection()` and created the R script `check_date()` in `/R` folder. Here is my R script for `check_date()`.

```
check_date <- function(date){
	tryCatch({as.Date(date); TRUE},
					 error = function(err) FALSE)
}
```

Before I added the helper function to `gday.R`, I tested the function by creating R script in the sub-folder `/tests/testthat` and run `devtools::test()` in the `R Console`. For example:
```
test_that("The date 20141-1-17 is not a valid date", {
	expect_false(check_date("20141-1-17"))
})
```

I also added the context to the top of the test R script as
```
context("Check helper function check_date")
```
so that everytime I run `devtools::test()`, I will know which context it is testing. After the tests are passed, I add the arguments to `gday.R`.
```
	if(check_date(date) == FALSE) {
		stop('Sorry, the date is invalid.')
	}
```

I did not use `assertthat` as suggested in the [Homework](http://stat545-ubc.github.io/hw10_package.html), since `if()` and `stop()` have more informative return. Then I also add tests about `gday()`. 

Sometimes you need to clean the objects created in `Environment` before `Build & Reload`. As you run those R scripts, objects with the same names will be created in the `Environment`. Thus R sometimes gets confused about whether you are testing a function in a package or your are using the function that you created in the `Environment`.

#### Add data in the package
I did spend sometime to figure out this part. I first read [How to add date to your package](http://r-pkgs.had.co.nz/data.html) and get to know that the data could be just put into the subdirectory `/data`, the data format in this folder is recommended to be `.rdata`, and that if there are some raw data, it could be put in the subdirectoy `/data-raw` with the R scripts that clean and reshape the data. I also followed the [R scripts in data-raw](https://github.com/hadley/babynames/blob/master/data-raw/applicants.R) of [babynames Package](https://github.com/hadley/babynames) from [Hadley Wickham's Github](https://github.com/hadley) to create data in `.rdata`  format like `save(arena_team, file = "data/arena_team.rdata")`.

However, how should I read `arena_team.rdata`? I consulted the [gapminder package](https://github.com/jennybc/gapminder) as suggested by Bernhard. The R scripts seem to just directly call the name of the data. There is no read-data argument such as `read.sdata`. I also tried it this way, and it worked. The only possible problem is when you `Build & Reload`, there would be warning messages saying that there is no documentation for the data. Then I added the documentation for `arena_team` following [R script](https://github.com/jennybc/gapminder/blob/master/R/country_colors.R) in [gapminder package](https://github.com/jennybc/gapminder).

#### Helper function `check_team()`
This helper function also serves the purpose to test the whether I can load the data. The R script is like this:

```
check_team <- function(team){
	any(grepl(team, arena_team$team, ignore.case = TRUE))
}

```

#### Function `gday()`
After all the helper functions are done with no unexpected error reported, the `gday()` function finally looks likes this:
```
gday <- function(team = "canucks", date = Sys.Date()) {
	if(internet_connection() == FALSE) {
		stop('Sorry, unable to connect to the Internet.
				    Please check your internet connection.')
	}
	if(check_date(date) == FALSE) {
		stop('Sorry, the date is invalid.')
	}
	if(check_team(team) == FALSE) {
		stop('Sorry, the city does not have an NHL team or the name has a typo.')
	}
	url <- paste0("http://live.nhle.com/GameData/GCScoreboard/", date, ".jsonp")
	grepl(team, RCurl::getURL(url), ignore.case=TRUE)
}

```

#### Function `scores()`
The `scores()` function finally looks like this:

```
scores <- function(date = Sys.Date()){
	if(internet_connection() == FALSE) {
		stop('Sorry, unable to connect to the Internet.
				    Please check your internet connection.')
	}
	if(check_date(date) == FALSE) {
		stop('Sorry, the date is invalid.')
	}
	url <- paste0("http://live.nhle.com/GameData/GCScoreboard/", date, ".jsonp")
	raw <- RCurl::getURL(url)
	json <- gsub('([a-zA-Z_0-9\\.]*\\()|(\\);?$)', "", raw, perl = TRUE)
	data <- jsonlite::fromJSON(json)$games
	with(data,
			 data.frame(home = paste(htn, htcommon),
			 					 away = paste(atn, atcommon),
			 					 home_score = hts,
			 					 away_score = ats))
}

```

#### Test the package
I created several test R scripts in the sub-directory `/tests/testthat`, each with a different context. I put this section in the end, but it does not mean that this is the last thing to do. I tested interactively the creation of functions.

## Reflections
The homework provides a great hands-on activity of building an R package that for students not in STAT department, they would rarely have the opportunity to experience. With the helpful TAs, I resolved all the problems encountered during the process. A lots of useful resources are provided in the [package page](http://stat545-ubc.github.io/packages00_index.html) that I could dig deeper later on. 


